---
title: "Week 32 - Ferris Wheels"
output: html_document
author: "Bear Jordan"
date: "2022-08-09"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r library, include=FALSE}
library(dplyr)
library(tidyr)
library(ggplot2)
library(forcats)
library(svglite)

theme_set(theme_minimal())
```

## Ferris Wheels

For this week, I want to ask did the development of ferris wheels in one country effect the development of ferris wheels in another country?

```{r load data}
wheels <- read.csv("wheels.csv")
head(wheels)
```


## Data Selection and EDA

Here is the data I plan to use for the figure.

```{r figure_data}
figure_data <- wheels %>% 
  select(country, opened, name, seating_capacity, status) %>% 
  drop_na()
head(figure_data)
```
```{r fix_date}
figure_data$opened <- as.Date(figure_data$opened)
figure_data$country <- as.factor(figure_data$country)
```


As a sanity check, I expect the capactity to increase through time.

```{r hyp_test}
plot(x=figure_data$opened,
     y=figure_data$seating_capacity,
     main="Does seating capacity increase?",
     xlab="",
     ylab="Seating Capacity",
     las=1)
```

Interesting, I wonder what happened in the earlier years.

```{r hyp_follow_up}
figure_data %>% 
  filter(opened<"1910-01-01") %>% 
  filter(seating_capacity>750)
```

The Great Wheel sounds wild---it was built in London after the sucess of the Chicago Ferris Wheel. No big horror stories it looks like though.

## Figure Construction

```{r plot 1}
ggplot(aes(x=opened, y=country, size=seating_capacity), data=figure_data) +
  geom_point(alpha=.3, stroke=.8) +
  geom_line(size=.4)
```

Well, this won't do. I want to sort these by the length of the line. And there is way to much empty space. I will try filtering the data again to only focus on things past 1950. I am just going to select countries with more than one ferris wheel too.

```{r new_data}
print(unique(figure_data$country))
country_filter <- figure_data %>% 
  count(country) %>% 
  filter(n!=1) %>% 
  select(country)

country_filter <- country_filter$country

figure_data <- figure_data %>% 
  filter(opened>"1955-01-01") %>% 
  filter(country %in% country_filter)
```

Okay, try this again.

```{r plot 2}
ggplot(aes(x=opened, y=country, size=seating_capacity), data=figure_data) +
  geom_point(alpha=.3, stroke=.8) +
  geom_line(size=.4)
```

Better maybe I will sort decending by the most recent ferris wheel.

```{r order}
plot_order <- figure_data %>% 
  group_by(country) %>% 
  top_n(1, opened) %>% 
  arrange(opened)

plot_order <- plot_order$country

figure_data$country <- factor(figure_data$country, levels = plot_order)
```

```{r plot 3}
ggplot(aes(x=opened, y=country, size=seating_capacity, color=status), data=figure_data) +
  geom_point(alpha=.3, stroke=.8) +
  geom_line(size=.4)
```

Okay, it seems crazy... does japan have one being built?

```{r japan}
wheels %>% 
  filter(country=="Japan") %>% 
  arrange(opened)
```

I want to clean up the status codes

```{r status}
figure_data <- figure_data %>% 
  filter(status %in% c("Defunct", "Operating", "Under Construction"))
```

```{r plot 4}
ggplot(aes(x=opened, y=country, size=seating_capacity, color=status), data=figure_data) +
  geom_point(alpha=.3, stroke=.8) +
  geom_line(size=.4)
```

Okay, why doesn't Malaysia have a line?

```{r malaysia}
figure_data %>% 
  filter(country=="Malaysia") %>% 
  ggplot(aes(x=opened, y=country, color=status)) +
      geom_point() +
      geom_line()
```

Okay, it looks like I need to place the color in the point.

```{r plot 5}
ggplot(aes(x=opened, y=country), data=figure_data) +
  geom_line(size=.4) +
  geom_point(alpha=1, stroke=.5, shape=21, color="black", aes(x=opened, y=country, size=seating_capacity), data=figure_data) +
  geom_point(alpha=.3, stroke=0, aes(color=status, x=opened, y=country, color=status, size=seating_capacity), data=figure_data) +
  theme(aspect.ratio=1) +
  ylab("") +
  xlab("") +
  scale_size_area(max_size=10)
```

```{r final plot}
final_plot <- ggplot(aes(x=opened, y=country), data=figure_data) +
  geom_line(size=.4) +
  geom_jitter(alpha=.3, stroke=.4, height=.12, aes(color=status, x=opened, y=country, size=seating_capacity), data=figure_data) +
  theme(aspect.ratio=1) +
  ylab("") +
  xlab("") +
  scale_size_area(max_size=10)
```

```{r export}
ggsave(file="final_plot.svg", plot=final_plot)
```

Awesome, there it is! I should include a label showing japan has 64 seats in the small ~1965 ferris wheel. And the ~2017 US one has a max of 1440.

What happened to Malaysia's recently defunct one?